---
title: "R Notebook"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "~/Dropbox/Datafiles/covid/WW_runs/Dec21OmicronSTW")
setwd("~/Dropbox/Datafiles/covid/WW_runs/Dec21OmicronSTW")
```

Read in EA grid ref etc
This file comes from something I got earlier in the year, misses geo info on a lot of sites, esp network

```{r}
EA.samples <- read.csv("~/Dropbox/Datafiles/covid/WW_runs/EA_samples/20210824_EA_WW_COVID_ASSURED.csv")
EA.lookup <- EA.samples[match(unique(EA.samples$sample_site_code),EA.samples$sample_site_code),c("sample_city_name", "sample_site_name", "sample_site_code", "ons_site_name",  "grid_reference_east", "grid_reference_north","grab_compo_ind")]
```

Will probably need to export EA.lookup and add in missing grid refs and missing sites added since

See how well this merges with STW list from a Dec run

```{r}
Dec17.run <- read.csv("~/Dropbox/Datafiles/covid/WW_runs/Dec21OmicronSTW/211217STW/211217_LIVERPOOL_UNI_RNA_PLATE_SURVEILLANCE.csv")
Dec17.run <- Dec17.run[nchar(Dec17.run$sample_site_code)>1,]
table(Dec17.run$sample_site_code %in% EA.lookup$sample_site_code)
#Dec17.run[!Dec17.run$sample_site_code %in% EA.lookup$sample_site_code,]

Dec17.run.wsite <- merge(Dec17.run,EA.lookup[,c("sample_site_code","grid_reference_east","grid_reference_north","grab_compo_ind")],by="sample_site_code",all.x=TRUE)
```

half a dozen or so sites not on the EA lookup, 26 missing grid refs (15 unique)

```{r}
#install.packages(c("rgdal","spdplyr","geojsonio","rmapshaper"),dependencies=TRUE)
#library(rgdal)
#library(spdplyr)
#library(geojsonio)
#library(rmapshaper)# Load UK Local Authorities' shapefile into R

#uk_la <- readOGR(dsn = "infuse_dist_lyr_2011_clipped", layer = "infuse_dist_lyr_2011_clipped")# Convert to Longitude / Latitude with WGS84 Coordinate System
#wgs84 = '+proj=longlat +datum=WGS84'
#uk_la_trans <- spTransform(uk_la, CRS(wgs84))# Convert from Spatial Dataframe to GeoJSON
#uk_la_trans_json <- geojson_json(uk_la_trans)# Simplify the polygons
#uk_la_trans_sim <- ms_simplify(uk_la_trans_json)# Save as GeoJSON file on the file system.
#geojson_write(uk_la_trans_sim, file = "uk_la.geojson")
```

try using simpler code
from https://bookdown.org/yann_ryan/r-for-newspaper-data/mapping-with-r-geocode-and-map-the-british-librarys-newspaper-collection.html
```{r}
library(ggplot2)
library(maps)
library(mapdata)

worldmap <- map_data('world')
UKmap <- map_data('worldHires',region="UK")
GBmap <- map_data('worldHires',region=c('UK', 'Ireland', 'Isle of Man','Isle of Wight', 'Wales:Anglesey'))

england.plot <- ggplot() + 
  geom_polygon(data = GBmap, 
               aes(x = long, y = lat, group = group), 
               fill = 'gray90', color = 'black') + 
  coord_fixed(ratio = 1.3, xlim = c(-6.5,3), ylim = c(50, 55.5)) + 
  theme_void()
england.plot
#add points to this
```

Need to convert eastings/northings to lat/long

```{r}
#install.packages("remotes")
#remotes::install_github("tayoso2/sfXtra")
library(sfXtra)

tst <- east_north_to_long_lat(431842, 430539, crs = 27700)

tst <- east_north_to_long_lat(Dec17.run.wsite$grid_reference_east,Dec17.run.wsite$grid_reference_north,crs=27700)

Dec17.run.wsite$longitude <- tst[,1]
Dec17.run.wsite$latitude <- tst[,2]
```

plot some points

```{r}
england.plot + geom_point(data = Dec17.run.wsite[!is.na(Dec17.run.wsite$longitude),], 
             aes(x = as.numeric(longitude), 
                 y = as.numeric(latitude)),fill="white",colour="black",shape=21,size=3)
```

Next step is link with Omicron frequency and fill with colour gradient

```{r}
setwd("~/Dropbox/Datafiles/covid/WW_runs/Dec21OmicronSTW")
load(file="acgt.muse.17df.Rdata")
#acgt.muse.17df <- acgt.muse.df
```

Add in omicron frequency
```{r}
Dec17.run.wfreq <- merge(Dec17.run.wsite,acgt.muse.df,by.x="sequencing_sample_id",by.y="sample",all.x=TRUE)
Dec17.run.wfreq$omicron.mu[Dec17.run.wfreq$omicron.mu == -1] <- NA
```

```{r}
tst <- england.plot + geom_point(data = Dec17.run.wfreq[!is.na(Dec17.run.wfreq$longitude) ,], 
             aes(x = as.numeric(longitude), 
                 y = as.numeric(latitude),fill=omicron.mu),colour="black",shape=21,size=3) 
tst +  scale_color_gradient(low = "white", high = "red",na.value = NA,aesthetics = "fill")
```

#only run once
```{r, eval=FALSE} 
write.csv(Dec17.run.wfreq,file="Dec17.run.wfreq.csv")
write.csv(EA.lookup,file="EA.lookup.csv") 
```

Export data and give an easy way to plot
Just need to 
```{r}

#essentially this is created by merging 
#(i) the spreadsheet going to DHSC with the sequencing ids, 
#(ii) the lookup table with longitude and latitude for each site (better for STWs than network) 
#(iii) some measure of omicron frequency per sample


Dec17.run.wfreq <- read.csv(file="Dec17.run.wfreq.csv")

#do the plots
library(ggplot2)
library(maps)
library(mapdata)

worldmap <- map_data('world')
UKmap <- map_data('worldHires',region="UK")
GBmap <- map_data('worldHires',region=c('UK', 'Ireland', 'Isle of Man','Isle of Wight', 'Wales:Anglesey'))

england.plot <- ggplot() + 
  geom_polygon(data = GBmap, 
               aes(x = long, y = lat, group = group), 
               fill = 'gray90', color = 'black') + 
  coord_fixed(ratio = 1.3, xlim = c(-6.5,3), ylim = c(50, 55.5)) + 
  theme_void()

england.plot + geom_point(data = Dec17.run.wfreq[!is.na(Dec17.run.wfreq$longitude) ,], 
             aes(x = as.numeric(longitude), 
                 y = as.numeric(latitude),fill=omicron.mu),colour="black",shape=21,size=3) +
             scale_color_gradient(low = "white", high = "red",na.value = NA,aesthetics = "fill")
```

